### 2022-10-27 manuscript.Rmd
Add new citations

  Number of words: 212 (abstract), 647 (introduction), 1,486 (discussion)

@@

The cognitive impenetrability hypothesis is challenged by the alternative view that perception dynamically interacts with different cognitive aspects from early on [@ahissar2004; @churchland1994; @clark2013; @lupyan2015; @lupyan2020; @yuille2006].
On the behavioral level, this is supported by psychological studies that found differences in ratings, detection rates, or reaction times for visual stimuli depending on their emotional [e.g., @phelps2016], motivational [e.g., @balcetis2010], linguistic [e.g., @boutonnet2015] or semantic [e.g., @gauthier2003] content.
However, some studies have received legitimate criticism for confounding tentative high-level effects with low-level differences between conditions, as well as for not being able to discern perceptual from post-perceptual (e.g., memory-related) effects [@firestone2016].

On the physiological level, event-related potentials (ERPs) measured from the human EEG can mitigate most of these concerns:
Their excellent temporal resolution makes it possible to probe how early one can detect influences of high-level (e.g., semantic) information.
To this end, participants in previous studies by our lab and others learned to associate visual objects with different amounts of semantic information.
After learning, an orthogonal task was used to compare ERPs in response to objects learned with versus without semantic information.
This revealed differences not only in late ERP components associated with semantic processing (i.e., the N400 component) but also in the visual P1 component [@abdelrahman2008; @maier2019; @samaha2018; @weller2019].
The early peak of this component and its source in the occipital cortex [@abdelrahman2008] point to an immediate effect of semantic knowledge on visual perception.
It is less clear, however, if this tentative top-down effect of semantic cognition on perception acts in an online fashion (i.e., directly modulating perceptual processing), or more indirectly, by altering stored visual representations over the course of learning, which would then get reactivated once the object is reencountered later on.

@@

Both the N170 and the P1 components therefore seem to be sensitive to the semantic meaningfulness of visual objects.
The finding that these two components were modulated in different phases of our experimental design suggests that they reflect different aspects of top-down processing with different time courses and neuroanatomical implementations.
The time course of the N170 is consistent with a top-down influence of (non-visual) areas in the prefrontal and temporal-parietal cortices on visual areas, whereas modulations of the P1 component seem to reflect recurrent processing within the visual system [@wyatte2014].
Here we could show that the former pathway seems to be able to convey semantic information instantaneously (i.e., within the same trial), whereas the latter pathway seems to take at least one additional encounter with the visual object to emerge.
While the limited spatial resolution of the EEG precludes precise localization, there is converging fMRI evidence showing that semantic information can feed back into higher-level areas in the lateral occipital cortex (LOC) and early retinotopic cortex [areas V1, V2, and V3\; @hsieh2010], consistent with the neural generators of the N170 and P1 components in the ERP.

@@

A theoretical framework that would explicitly predict the observed P1 and N170 effects in our study is lacking at present.
However, the effects are consistent with the reverse hierarchy theory [@ahissar2004], which posits that objects first enter visual consciousness at an abstract, conceptual level. Once this initial "vision at a glance" has taken place, feedback connections to earlier layers of the visual system are being accessed to extract the relevant lower-level features ("vision with scrutiny").
This reverse trajectory down the visual hierarchy may explain (a) the semantically induced changes to the fMRI signal in LOC and retinotopic cortex [@hsieh2010] as well as (b) the modulations of early visual ERP components observed in the present study and others [e.g., @abdelrahman2008; @maier2014; @samaha2018].
An important role of top-down mechanisms for object recognition is also posited by theories of predictive coding and Bayesian inference [e.g., @yuille2006; @xu2007; @clark2013; @panichello2013; @lupyan2015].
Despite the theoretical advances, detailed descriptions of these top-down effects at the algorithmic and implementational levels remain a challenge for future work.

>>>>

  Number of words: 215 (abstract), 641 (introduction), 1,490 (discussion)

@@

The cognitive impenetrability hypothesis is challenged by the alternative view that perception dynamically interacts with different cognitive aspects from early on [@ahissar2004; @churchland1994; @clark2013; @lupyan2015; @lupyan2020; @thierry2016; @yuille2006].
On the behavioral level, this is supported by psychological studies that found differences in ratings, detection rates, or reaction times for visual stimuli depending on their emotional [e.g., @phelps2016], motivational [e.g., @balcetis2010], linguistic [e.g., @slivac2021] or semantic [e.g., @gauthier2003] content.
However, some studies have received legitimate criticism for confounding tentative high-level effects with low-level differences between conditions, as well as for not being able to discern perceptual from post-perceptual (e.g., memory-related) effects [@firestone2016].

On the physiological level, event-related potentials (ERPs) measured from the human EEG can mitigate most of these concerns:
Their excellent temporal resolution makes it possible to probe how early one can detect influences of high-level (e.g., semantic) information [@athanasopoulos2020].
To this end, participants in previous studies by our lab and others learned to associate visual objects with different amounts of semantic information.
After learning, an orthogonal task was used to compare ERPs in response to objects learned with versus without semantic information.
This revealed differences not only in late ERP components associated with semantic processing (i.e., the N400 component) but also in the visual P1 component [@abdelrahman2008; @maier2019; @samaha2018; @weller2019].
The early peak of this component and its source in the occipital cortex [@abdelrahman2008] point to an immediate effect of semantic knowledge on visual perception.
It is less clear, however, if this tentative top-down effect of semantic cognition on perception acts in an online fashion (i.e., directly modulating perceptual processing), or more indirectly, by altering stored visual representations over the course of learning, which would then get reactivated once the object is reencountered later on.

@@

Both the N170 and the P1 components therefore seem to be sensitive to the semantic meaningfulness of visual objects.
The finding that these two components were modulated in different phases of our experimental design suggests that they reflect different aspects of top-down processing with different time courses and neuroanatomical implementations.
The time course of the N170 is consistent with a top-down influence of (non-visual) areas in the prefrontal and temporal-parietal cortices on visual areas, whereas modulations of the P1 component seem to reflect recurrent processing within the visual system [@wyatte2014].
Here we could show that the former pathway seems to be able to convey semantic information instantaneously (i.e., within the same trial), whereas the latter pathway seems to take at least one additional encounter with the visual object to emerge.
While the limited spatial resolution of the EEG precludes precise localization, there is converging fMRI and psychophysical evidence that semantic information can feed back into higher-level areas in the lateral occipital cortex (LOC) as well as early retinotopic cortex [areas V1, V2, and V3\; @hsieh2010; @clarke2016; @teufel2018], consistent with the neural generators of the N170 and P1 components in the ERP.

@@

A theoretical framework that would explicitly predict the observed P1 and N170 effects in our study is lacking at present.
However, the effects are consistent with the reverse hierarchy theory [@ahissar2004], which posits that objects first enter visual consciousness at an abstract, conceptual level. Once this initial "vision at a glance" has taken place, feedback connections to earlier layers of the visual system are being accessed to extract the relevant lower-level features ("vision with scrutiny").
This reverse trajectory down the visual hierarchy may explain (a) the semantically induced changes to the fMRI signal in LOC and retinotopic cortex [e.g., @hsieh2010] as well as (b) the modulations of early visual ERP components observed in the present study and others [e.g., @abdelrahman2008; @maier2014; @samaha2018].
An important role of top-down mechanisms for object recognition is also posited by theories of predictive coding and Bayesian inference [e.g., @yuille2006; @xu2007; @clark2013; @panichello2013; @lupyan2015].
Despite the theoretical advances, detailed descriptions of these top-down effects at the algorithmic and implementational levels remain a challenge for future work.
